Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2016-12-16 ms_pre_post.Rmd
Edit up to second to last paragraph of Introduction

@@

Verbal theory
----------------------------------------------

A recent genetically-explicit individual-based model showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance, and resulting joint evolution of these reproductive strategies, occurred only under highly constrained conditions [@Duthie]. Specifically, such evolution occurred when direct costs of polyandry were small, when very few males were available for a female's initial mate choice but many were available for additional mate choice(s), or when polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. In the absence of these conditions, increasing degrees of polyandry tended to increase the overall degree of inbreeding that females experienced, and hence increase inbreeding depression in offspring fitness. This occurred because polyandrous females that had evolved to avoid inbreeding had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance was consequently restricted.

However, while @Duthie considered simultaneous evolution of polyandry and pre-copulatory inbreeding avoidance, they did not also allow simultaneous evolution of post-copulatory inbreeding avoidance, and assumed that polyandrous females had no post-copulatory control over fertilisation. If post-copulatory inbreeding avoidance could also evolve, polyandrous females could further bias paternity towards unrelated males within their set of chosen mates. This process might reduce the cost of polyandry stemming from accumulation of related mates, potentially facilitating evolution of polyandry to avoid inbreeding [@Zeh1997], and driving further consequent evolution of pre-copulatory or post-copulatory mate choice strategies. Therefore, while pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions [@Duthie], the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin. The long-term evolutionary dynamics of pre-copulatory and post-copulatory inbreeding avoidance then become unclear. Strong inbreeding depression might drive initial evolution of both pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. However, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance redundant, or vice versa, then only one strategy might be maintained in the long-term.

Previous models implicitly or explicitly considered the fate of a rare allele underlying pre-copulatory inbreeding avoidance in a population initially fixed for random mating [e.g., @Parker1979; @Parker2006; @Duthie; @Duthie2016a]. Such models thereby isolate the invasion fitness of this single strategy. However, when both pre-copulatory and post-copulatory strategies can affect realised inbreeding, it cannot be assumed that both strategies will invade simultaneously, nor that invasion fitness of one strategy will be independent of the other. For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect of such invading alleles could be weak or negligible.

>>>>

A recent genetically-explicit individual-based model examined the conditions under which polyandry is predicted to evolve slely due to selection stemming from pre-copulatory inbreeding avoidance [@Duthie]. Evolution of polyandry occurred only under highly restricted conditions, requiring that direct costs were small, that very few males were available for a female's initial mate chioce but many were available for additional mate choice(s), or that polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. Without these conditions, increasing polyandry tended to increase rather than decrease the overall degree of inbreeding that females experienced, and hence reduced offspring fitness. This occurred because, once pre-copulatory inbreeding avoidance evolved, polyandrous females had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance, and hence the resulting degree of inbreeding avoidance, was consequently restricted [@Duthie].

<!--- This is getting awfully narrow in terms of theory; postcopulatory *might* happen given a fairly specific set of conditions concerning evolution of pre-copulatory inbreeding avoidance and polyandry. Are we confident these conditions are general? --->

<!--- I'm not suggesting that things need to be changed (probably too late for that), just that, to me, this reads as the theoretical equivalent of a very system-specific question being address, not like a question that has broad implications across different systems --->

However, if post-copulatory inbreeding avoidance could also evolve, polyandrous females could further bias paternity towards unrelated males within their set of chosen mates. Post-copulatory inbreeding avoidance might consequently reduce the cost of polyandry stemming from the accumulation of related mates, potentially facilitating evolution of polyandry to avoid inbreeding under broader conditions, and driving further consequent evolution of pre-copulatory or post-copulatory mate choice strategies. However, once both pre-copulatory and post-copulatory inbreeding avoidance, and resulting polyandry, can evolve, the long-term dynamics of these reproductive strategies become difficult to conceptualise and predict. Strong inbreeding depression might drive initial evolution of both pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry. However, if evolution of polyandry and post-copulatory inbreeding avoidance renders pre-copulatory inbreeding avoidance redundant, or vice versa, then only one inbreeding avoidance strategy might be maintained in the long-term.

The few previous models that considered the evolution of inbreeding avoidance through mate choice (as opposed to dispersal) have implicitly or explicitly considered the fate of a rare allele underlying pre-copulatory inbreeding avoidance in a population initially fixed for random mating [e.g., @Parker1979; @Parker2006; @Duthie; @Duthie2016a]. Such models are useful for isolating the invasion fitness of this single strategy. However, when both pre-copulatory and post-copulatory strategies can affect realised inbreeding, it cannot be assumed that both strategies will simultaneously invade a randomly mating population, nor that the invasion fitness of one strategy will be independent of the existence of invasion fitness of the other strategy. For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then new alleles underlying polyandry and post-copulatory inbreeding avoidance might be precluded from invading a population because the phenotypic effect of such alleles on the degree of inbreeding could be negligible. Conversely, fixation of alleles underlying polyandry and post-copulatory inbreeding avodiance might eliminate positive selection on alleles underlying pre-copulatory inbreeding avoidance. New theory, guided by modelling that evaluates invasion dynamics given appropriate sets of evolutionary constraints, is therefore needed.

@@
